---
title: "Results"
author: "Ella Orme"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
setwd("/Users/ellaorme/Documents/resNMTF/")
source("Investigation/toy_data_gen.r")
setwd("/Users/ellaorme/Documents/resNMTF/SimStudy/RunSim/")
invisible(source("Functions/extra_funcs.r"))
invisible(source("main.r"))
source("Functions/evaluation_funcs.r")
row_dims <- rep(200, 2)
# col_dims <- rep(250, 2)
col_dims <- c(150, 250)
no_bicl <- 3
noise <- 5
val <- 200
n_views <- length(row_dims)
phi_mat <- matrix(0, n_views, n_views)
phi_mat[1,2] <- 1
psi_mat <- matrix(0, n_views, n_views)
psi_mat[1,2] <- 1
#differing no of biclusters
diff_bicl <- c(3, 4)
overlap <- 0.1
non_exh <- 0.9
phi_vec <- seq(0,1000,200)
folder_path <- "Investigation/toy_2v3b5n/data"
setwd("/Users/ellaorme/Documents/resNMTF")
```

We consider a base scenario of `r n_views` views (both `r row_dims[1]`x`r col_dims[1]`) with `r no_bicl` shared row clusters across views. A noise level of `r noise` is used. 

# Example 1
## Noisy views
(a) We first consider the results when the second view is pure noise.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_1a")
ex_1a <- multi_view_toy(file_path,row_dims, col_dims,
             no_bicl, noise, row_e = c(1,0), col_e = c(1,0), seed=4)
res_1a <- toy_example(ex_1a, file_path, phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```

(b) When all views are noise.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_1b")
ex_1b <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise, row_e = 0, col_e = 0, seed=20)
res_1b <- toy_example(ex_1b, file_path, phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```

# Example 2
## Shared column clusters
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_2")
ex_2 <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise, col_same_shuffle = TRUE, seed=100)      
res_2 <- toy_example(ex_2, file_path, phi_vec=phi_vec, phi=phi_mat, psi=psi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```

# Example 3
## Differing number of biclusters
View 2 now has `r diff_bicl[2]` biclusters. This assumes no shared clustering in rows or columns. 
```{r, echo=FALSE, warning=FALSE}
file_path <- paste0(folder_path,"/ex_3")
ex_3 <- multi_view_toy(file_path, row_dims, col_dims,
             diff_bicl, noise,  seed=5)
res_3 <- restMultiNMTF_run(Xinput = ex_3$data_views)
all_results <- evaluate_simulation_comp(res_3$row_clusters,
                 res_3$col_clusters,
            ex_3$truth_rows, ex_3$truth_cols, ex_3$data_views)
row_filename <- paste0(file_path,  "/row_clusts.xlsx")
col_filename <- paste0(file_path, "/col_clusts.xlsx")
openxlsx::write.xlsx(res_3$row_clusters, file = row_filename)
openxlsx::write.xlsx(res_3$col_clusters, file = col_filename)
f_vec <- (mean(all_results$F_score))
rec_vec <- (mean(all_results$Rec))
rel_vec <- (mean(all_results$Rel))
print(paste0("This gives f-score ", f_vec,
         ", recovery score ", rec_vec,
          " and relevance score ", rel_vec,"."))
summary <- data.frame("phi" = phi_vec, "f_score" = f_vec,
         "rec" =  rec_vec, "rel" =  rel_vec, "BiS" = res_3$Sil_score)
write.csv(summary, file = paste0(file_path, "/summary.csv"))
read.csv(paste0(file_path,"/summary.csv"))
```


# Example 4
## Overlapping biclusters
(a) Overlapping columns, `r overlap*100`% of a bicluster overlaps with the one beside it (so bicluster 2 overlaps with bicluster 1 and 3).
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_4a")
ex_4a <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise,  col_o = overlap,  seed=6)
res_4a <- toy_example(ex_4a,file_path, phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```

(b) Overlapping rows, `r overlap*100`% overlap as above.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_4b")
ex_4b <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise, row_o = overlap,  seed=7)
res_4b <- toy_example(ex_4b, file_path, phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```

# Example 5
## Non- exhaustivitiy

(a) Non-exhaustive columns, `r non_exh*100`% of columns do not belong to any bicluster.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_5a")
ex_5a <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise,  col_e = non_exh,  seed=6)
res_5a <- toy_example(ex_5a, file_path, phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```
(b) Non-exhaustive rows, `r non_exh*100`% of rows do not belong to any bicluster.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_5b")
ex_5b <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise, row_e = non_exh,  seed=7)
res_5b <- toy_example(ex_5b,file_path,  phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```

# Examples 6
## Combinations
(a) Non-exhaustive and non-exclusive columns, `r non_exh*100`% of columns do not belong to any bicluster and `r overlap*100`% overlap.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_6a")
ex_6a <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise,  col_e = non_exh, col_o = overlap, seed=6)
res_6a <- toy_example(ex_6a,file_path,  phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```
(b) Non-exhaustive and non-exlusive rows, `r non_exh*100`% of rows do not belong to any bicluster and `r overlap*100`% overlap.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_6b")
ex_6b <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise, row_e = non_exh, row_o = overlap, seed=7)
res_6b <- toy_example(ex_6b,file_path,  phi_vec=phi_vec, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
```
(c) Non-exhaustivitiy and non-exclusitivity for both rows and columns, both at `r overlap*100`%.
```{r, echo=FALSE}
file_path <- paste0(folder_path,"/ex_6c")
ex_6c <- multi_view_toy(file_path, row_dims, col_dims,
             no_bicl, noise, row_e = non_exh, row_o = 0.2,
            col_e = non_exh, col_o = 0.2, seed=7)
res_6c <- toy_example(ex_6c,file_path,  phi_vec=200, phi=phi_mat)
read.csv(paste0(file_path,"/summary.csv"))
row_clusters <- res_6c$row_clusters[[1]]
```

(d)
```{r, echo=FALSE}
rows <- import_matrix("Investigation/toy_2v3b5n/data/ex_6c/200_row_clusts.xlsx")
cols <- import_matrix("Investigation/toy_2v3b5n/data/ex_6c/200_col_clusts.xlsx")

# re order row cluster
row_order <- c()
# row cluster 1, no intersection with cluster 3
row_order <- c(row_order, (1:200)[(rows[[1]][,1]==1)&(rows[[1]][,3]==0)])
# row cluster 1, intersection with cluster 3
row_order <- c(row_order,
          (1:200)[(rows[[1]][,1]==1)&(rows[[1]][,3]==1)])
# row cluster 3, no intersection with cluster 1 or 2
row_order <- c(row_order, 
          (1:200)[(rows[[1]][,3]==1)&(rows[[1]][,1]==0)&(rows[[1]][,2]==0)])
# row cluster 3, intersection with cluster 2
row_order <- c(row_order, 
          (1:200)[(rows[[1]][,3]==1)&(rows[[1]][,2]==1)])
# row cluster 2, no intersection with cluster 3
row_order <- c(row_order, 
          (1:200)[(rows[[1]][,2]==1)&(rows[[1]][,3]==0)])
# no row clusters
row_order <- c(row_order, 
          (1:200)[rowSums(rows[[1]])==0])

# re order clusters
col_order <- c()
# col cluster 1, no intersection with cluster 3
col_order <- c(col_order, (1:150)[(cols[[1]][,1]==1)&(cols[[1]][,3]==0)])
# row cluster 1, intersection with cluster 3
col_order <- c(col_order,
          (1:150)[(cols[[1]][,1]==1)&(cols[[1]][,3]==1)])
# row cluster 3, no intersection with cluster 1 or 2
col_order <- c(col_order, 
          (1:150)[(cols[[1]][,3]==1)&(cols[[1]][,1]==0)&(cols[[1]][,2]==0)])
# row cluster 3, intersection with cluster 2
col_order <- c(col_order, 
          (1:150)[(cols[[1]][,3]==1)&(cols[[1]][,2]==1)])
# row cluster 2, no intersection with cluster 3
col_order <- c(col_order, 
          (1:150)[(cols[[1]][,2]==1)&(cols[[1]][,3]==0)])   
# no row clusters
col_order <- c(col_order, 
          (1:150)[rowSums(cols[[1]])==0])
data <- import_matrix("/Users/ellaorme/Documents/resNMTF/Investigation/toy_2v3b5n/data/ex_6c/data.xlsx") 
pdf('Investigation/toy_2v3b5n/data/ex_6c/view1_biclustered.pdf')
image(data[[1]][row_order,unique(col_order)])
dev.off()  

col_order2 <- c()
# col cluster 1, no intersection with cluster 3
col_order2 <- c(col_order2, (1:250)[(cols[[2]][,1]==1)&(cols[[2]][,3]==0)])
# row cluster 1, intersection with cluster 3
col_order2 <- c(col_order2,
          (1:250)[(cols[[2]][,1]==1)&(cols[[2]][,3]==1)])
# row cluster 3, no intersection with cluster 1 or 2
col_order2 <- c(col_order2, 
          (1:250)[(cols[[2]][,3]==1)&(cols[[2]][,1]==0)&(cols[[2]][,2]==0)])
# row cluster 3, intersection with cluster 2
col_order2 <- c(col_order2, 
          (1:250)[(cols[[2]][,3]==1)&(cols[[2]][,2]==1)])
# row cluster 2, no intersection with cluster 3
col_order2 <- c(col_order2, 
          (1:250)[(cols[[2]][,2]==1)&(cols[[2]][,3]==0)])   
# no row clusters
col_order2 <- c(col_order2, 
          (1:250)[rowSums(cols[[2]])==0])
# dev.off()
pdf('Investigation/toy_2v3b5n/data/ex_6c/view2_biclustered.pdf')
image(data[[2]][row_order,col_order2])
dev.off()
pdf('Investigation/toy_2v3b5n/data/ex_6c/view2_shuffle.pdf')
image(data[[2]][sample(200),sample(250)])
dev.off()
pdf('Investigation/toy_2v3b5n/data/ex_6c/view1_shuffle.pdf')
image(data[[1]][sample(200),sample(150)])
dev.off()
```

```{r}
true_rows <- import_matrix("/Users/ellaorme/Documents/resNMTF/Investigation/toy_2v3b5n/data/ex_6c/true_rows.xlsx")
true_cols <- import_matrix("/Users/ellaorme/Documents/resNMTF/Investigation/toy_2v3b5n/data/ex_6c/true_cols.xlsx")
sum((cols[[2]][,1]==1)&(cols[[2]][,3]==1))
sum((true_cols[[2]][,1]==1)&(true_cols[[2]][,2]==1))
```