---
title: "Title"
author: "Author"
date: "Date"
output: 
  cleanrmd::html_document_clean:
    theme: stylize  # see cleanrmd::cleanrmd_themes()
                    # or set to NULL for dynamic theme picker (requires internet)
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("biclust")
library(biclust)
library(bisilhouette)
apply_biclust_method <- function(data, method, name,
                                 transposed = FALSE, number_clusts = 6) {
  row_clusters <- vector("list", length = length(data))
  col_clusters <- vector("list", length = length(data))
  for (i in seq_along(data)) {
    if (name == "Motif") {
      res <- quiet(biclust(data[[i]], method = method, number = number_clusts))
    } else {
      res <- suppressMessages(biclust(data[[i]], method = method))
    }
    if (anyNA(res@RowxNumber)) {
      row_clusters[[i]] <- matrix(0, nrow = nrow(data[[i]]), ncol = 1)
      col_clusters[[i]] <- matrix(0, nrow = ncol(data[[i]]), ncol = 1)
    } else {
      rows <- apply(res@RowxNumber, 2, as.numeric)
      cols <- apply(res@NumberxCol, 2, as.numeric)
      if (!is.matrix(cols)) {
        col_clusters[[i]] <- matrix(cols, nrow = length(cols), ncol = 1)
        row_clusters[[i]] <- matrix(rows, nrow = length(rows), ncol = 1)
      } else {
        col_clusters[[i]] <- t(cols)
        row_clusters[[i]] <- rows
      }
    }
  }
  if (transposed) {
    return(list(row_clusters = col_clusters, col_clusters = row_clusters))
  }
  return(list(row_clusters = row_clusters, col_clusters = col_clusters))
}
biclust_results <- function(data, labels, method, name,
                            transposed = FALSE, repeats = 5,
                            number_clusts = 6) {
  n_cols <- (length(data) + 1) * 4
  res_mat <- matrix(0, nrow = repeats, ncol = n_cols)
  for (i in seq_len(repeats)) {
    results <- apply_biclust_method(data, method, name,
                                    transposed, number_clusts)
    # return(results)
    if (transposed) { 
      jaccs <- all_jaccs(labels, results$row_clusters)
      bisils <- calc_euc_sils(lapply(data, t), results)
    } else {
      bisils <- calc_euc_sils(data, results)
      jaccs <- all_jaccs(labels, results$row_clusters)
    }
    res_mat[i, ] <- c(jaccs, bisils)
  }
  res_mat <- cbind(rep(name, repeats), res_mat)
  return(res_mat)
}

```
# Load the data

Data is terms by documents, 
Need to 
- check if any of the methods require additional parameters
- store results in a csv file
- getting bisil of 1 when all are in one cluster
```{r}
# no params needed
bcplaid_results <- biclust_results(three_data, docs_labs,
                                   method = BCPlaid, name = "Plaid",
                                   transposed = TRUE)
# for questionnaire data - so haven't used BCQuest
bcspectral_results <- biclust_results(three_data, docs_labs,
                                      method = BCSpectral, name = "spectral",
                                      transposed = TRUE)
bcplaid_results2 <- biclust_results(lapply(three_data, t), docs_labs,
                                    method = BCPlaid, name = "Plaid_t")
bcspectral_results2 <- biclust_results(lapply(three_data, t), docs_labs,
                                       method = BCSpectral, name = "spectral_t")

full_res <- rbind(bcplaid_results,
                  bcplaid_results2,
                  bcspectral_results,
                  bcspectral_results2)
write.csv(full_res, "RealData/3sources/biclust_results_term_doc.csv",
          row.names = FALSE)


```

```{r}
source("SimStudy/Functions/extra_funcs.r")
source("SimStudy/Functions/evaluation_funcs.r")


three_cancers <- import_matrix("RealData/single_cell", "data_processed")
labels <- read.csv("RealData/single_cell/true_labs.csv", row.names=1)

plaid_res <- biclust(three_cancers[[2]], method = "BCPlaid")
rows <- apply(plaid_res@RowxNumber, 2, as.numeric)
cols <- apply(plaid_res@NumberxCol, 2, as.numeric)
jaccard_row(labels, rows)

# no params needed
bcplaid_results <- biclust_results(three_data, docs_labs,
                                   method = BCPlaid, name = "Plaid",
                                   transposed = TRUE)
# for questionnaire data - so haven't used BCQuest
bcspectral_results <- biclust_results(three_data, docs_labs,
                                      method = BCSpectral, name = "spectral",
                                      transposed = TRUE)
bcplaid_results2 <- biclust_results(lapply(three_data, t), docs_labs,
                                    method = BCPlaid, name = "Plaid_t")
bcspectral_results2 <- biclust_results(lapply(three_data, t), docs_labs,
                                       method = BCSpectral, name = "spectral_t")

full_res <- rbind(bcplaid_results,
                  bcplaid_results2,
                  bcspectral_results,
                  bcspectral_results2)
write.csv(full_res, "RealData/3sources/biclust_results_term_doc.csv",
          row.names = FALSE)
```



## 3cancers results
```{r}
source("SimStudy/Functions/extra_funcs.r")
source("SimStudy/Functions/evaluation_funcs.r")


three_cancers <- import_matrix("RealData/3cancers", "tgca_3cancers")
labels <- read.csv("RealData/3cancers/tgca_labels.csv")
tc_data <- lapply(three_cancers, function(x) apply(x, 2, as.numeric))
tc_data <- lapply(tc_data, function(x) x[, 2:1242])

```

```{r}
library(biclust)
# no params needed
bcplaid_results <- biclust_results(tc_data, labels,
                                   method = BCPlaid, name = "Plaid",
                                   transposed = TRUE)

# for questionnaire data - so haven't used BCQuest
bcspectral_results <- biclust_results(tc_data, docs_labs,
                                      method = BCSpectral, name = "spectral",
                                      transposed = TRUE)bcplaid_results2 <- biclust_results(lapply(tc_data, t), docs_labs,
                                    method = BCPlaid, name = "Plaid_t")
bcspectral_results2 <- biclust_results(lapply(tc_data, t), docs_labs,
                                       method = BCSpectral, name = "spectral_t")

# biclust(three_cancers[[2]], method = BCPlaid)

full_res <- rbind(bcplaid_results,
                  bcspectral_results)
# write.csv(full_res, "RealData/3cancers/biclust_results_term_doc.csv",
#           row.names = FALSE)
best_res <- matrix(0, nrow = 2, ncol = ncol(full_res))
i <- 1
for (method in c("Plaid", "spectral")) {
  res <- full_res[full_res[, 1] == method, ]
  best_res[i, ] <- res[which.max(res[, 5]), ]
  i <- i + 1
}
# only keep non transposed results
write.csv(best_res[c(1, 3), ], "RealData/3cancers/biclust_results.csv",
          row.names = FALSE)

cols <- apply(mvrnorm(n = 20267, mu = rep(0, 3), Sigma = diag(3)) > 0, 2, as.numeric)
bisilhouette(t(tc_data[[1]]),
        labels,
        cols,
        method = "cosine",
      )
```

```{r}
k <- 1
filepath <- "RealData/3cancers"
transposed_data <- FALSE
for (t in 1:5) {
  row_clusts <- import_matrix(
    paste0(filepath, "/issvd_res/"),
    paste0(t - 1, "_row_clusts")
  )
  col_clusts <- import_matrix(
    paste0(filepath, "/issvd_res/"),
    paste0(t - 1, "_col_clusts")
  )
  set.seed(10 + t)
  if (transposed_data) {
    clust_res <- list(
      "row_clusters" = col_clusts,
      "col_clusters" = row_clusts
    )
  } else {
    clust_res <- list(
      "row_clusters" = row_clusts,
      "col_clusters" = col_clusts
    )
  }
  # bisils <- calc_all_sils(
  #   input_data,
  #   clust_res
  # )
  jaccs <- all_jaccs(labels, clust_res$col_clusters)
  print(jaccs)
}
# jaccs <- all_jaccs(labels, clust_res$row_clusters)
```

Now look at the GFA results
```{r}
gfa_res <- gfa_apply(lapply(tc_data, t), dim(labels)[2])
colSums(gfa_res$col_clusters[[2]])
jacc_res <- all_jaccs(labels, gfa_res$row_clusters)
bisils <- calc_euc_sils(lapply(tc_data, t), gfa_res)
bisils <- calc_all_sils(lapply(tc_data, t), gfa_res)
gfa_full_res <- c("GFA", jacc_res, bisils)

write.csv(rbind(gfa_full_res,
                best_res),
          "RealData/3cancers/biclust_results.csv",
          row.names = FALSE)
```
```{r}
#' @return TRUE if any columns/rows with only zeros, FALSE otherwise
test_cond <- function(data, attempt) {
  if (attempt == 1) {
    return(TRUE)
  }
  return(any(unlist(lapply(
    data,
    function(x) {
      any(colSums(x) == 0) | any(rowSums(x) == 0)
    }
  ))))
}

#' @title number of biclusters
#' @description determine number of biclusters
#' @param results results of apply_resnmtf/res_nmtf_inner
#' @noRd
#' @return number of biclusters
number_biclusters <- function(results) {
  return(sum(as.numeric(lapply(
    results$row_clusters,
    function(x) sum(colSums(x))
  ))))
}

#' @title shuffle view
#' @description shuffle view
#' @param data list of matrices to be shuffled
#' @param new_data list of shuffled matrices
#' @param dims dimensions of the view to be shuffled
#' @param dim dimensions of the first view
#' @param dim_1 dimensions of the first view
#' @param i index of the view to be shuffled
#' @param sample_rate rate at which to sample
#' @param row_samples list of rows to sample
#' @param col_samples list of cols to sample
#' @noRd
initial_shuffle <- function(
    data, new_data, dims, dim, i,
    sample_rate, row_samples, col_samples) {
  if ((dims[1]) == dim[1]) {
    row_samples[[i]] <- row_samples[[1]]
  } else {
    row_samples[[i]] <- sample(dims[1], (dims[1] * sample_rate))
  }
  if ((dims[2]) == dim[2]) {
    col_samples[[i]] <- col_samples[[1]]
  } else {
    col_samples[[i]] <- sample(dims[2], (dims[2] * sample_rate))
  }
  new_data[[i]] <- data[[i]][row_samples[[i]], col_samples[[i]]]
  return(list(
    "new_data" = new_data,
    "row_samples" = row_samples,
    "col_samples" = col_samples
  ))
}

#' @title check for empty matrix
#' @description check if matrix has any rows/columns with only zeros
#' @param data list of matrices to be checked
#' @param i index of the view to be checked
#' @noRd
#' @return TRUE if any rows/columns with only zeros, FALSE otherwise
check_empty <- function(data, i) {
  return(any(colSums(data[[i]]) == 0) || any(rowSums(data[[i]]) == 0))
}

#' @title sample view
#' @description sample view
#' @param data list of matrices to be shuffled
#' @param new_data list of shuffled matrices
#' @param dims dimensions of the view to be shuffled
#' @param dim_1 dimensions of the first view
#' @param i index of the view to be shuffled
#' @param row_samples list of rows to sample
#' @param col_samples list of cols to sample
#' @param sample_rate rate at which to sample
#' @noRd
#' @return list of shuffled matrices (new_data),
#'         rows (row_samples) and columns (col_samples)
sample_view <- function(data, i, new_data, dim_1,
                        row_samples, col_samples, sample_rate) {
  dims <- dim(data[[i]])
  shuffled <- initial_shuffle(
    data, new_data, dims, dim_1, i, sample_rate, row_samples, col_samples
  )
  new_data <- shuffled[["new_data"]]
  row_samples <- shuffled[["row_samples"]]
  col_samples <- shuffled[["col_samples"]]
  if (check_empty(new_data, i)) {
    zeros_cols <- colSums(new_data[[i]]) != 0
    zeros_rows <- rowSums(new_data[[i]]) != 0
    if ((dims[1]) == dim_1[1]) {
      for (p in 1:i) {
        row_samples[[p]] <- row_samples[[p]][zeros_rows]
      }
    } else {
      row_samples[[i]] <- row_samples[[i]][zeros_rows]
    }
    if ((dims[2]) == dim_1[2]) {
      for (p in 1:i) {
        col_samples[[p]] <- col_samples[[p]][zeros_cols]
      }
    } else {
      col_samples[[i]] <- col_samples[[i]][zeros_cols]
    }
    for (p in 1:i) {
      new_data[[p]] <- data[[p]][row_samples[[p]], col_samples[[p]]]
    }
  }
  return(list(
    "new_data" = new_data,
    "row_samples" = row_samples,
    "col_samples" = col_samples
  ))
}
```
```{r}
stability_repeat <- function(data, dim_1, 
                             n_views, sample_rate) {
  new_data <- vector(mode = "list", length = n_views)
  row_samples <- vector(mode = "list", length = n_views)
  col_samples <- vector(mode = "list", length = n_views)
  relevance <- matrix(0, nrow = n_views, ncol = k)
  # turn this into a function to be used with lapply
  attempt <- 1
  # need to check that
  while (test_cond(data, attempt)) {
    if (attempt == 20) {
      print("Unable to perform stability analysis due to sparsity of data.")
      return(list("stability_performed" = FALSE))
    }
    row_samples[[1]] <- sample(dim_1[1], (dim_1[1] * sample_rate))
    col_samples[[1]] <- sample(dim_1[2], (dim_1[2] * sample_rate))
    new_data[[1]] <- data[[1]][row_samples[[1]], col_samples[[1]]]
    if (check_empty(new_data, 1)) {
      zeros_cols <- colSums(new_data[[1]]) != 0
      zeros_rows <- rowSums(new_data[[1]]) != 0
      row_samples[[1]] <- row_samples[[1]][zeros_rows]
      col_samples[[1]] <- col_samples[[1]][zeros_cols]
      new_data[[1]] <- data[[1]][row_samples[[1]], col_samples[[1]]]
    }
    if (n_views > 1) {
      for (i in 2:n_views) {
        shuffled <- sample_view(
          data, i, new_data, dim_1, row_samples, col_samples, sample_rate
        )
        new_data <- shuffled[["new_data"]]
        row_samples <- shuffled[["row_samples"]]
        col_samples <- shuffled[["col_samples"]]
      }
    }
    return(new_data)
    attempt <- attempt + 1
  }
  }

res <- stability_repeat(input_data, dim(input_data[[1]]), 3, 0.95)

```
stability_repeat(c(), data, dim_1, k, phi, xi, psi, n_iters,
                             num_repeats, distance, n_views, sample_rate) 

## 3cancers results
```{r}
source("SimStudy/Functions/extra_funcs.r")
source("SimStudy/Functions/evaluation_funcs.r")


three_cancers <- import_matrix("RealData/3cancers/tgca_3cancers.xlsx")
labels <- read.csv("RealData/3cancers/tgca_3cancers_labels.csv")
tc_data <- lapply(three_cancers, function(x) apply(x, 2, as.numeric))
tc_data <- lapply(tc_data, function(x) x[, 2:1242])
```

```{r}
library(biclust)
# no params needed
bcplaid_results <- biclust_results(tc_data, labels,
                                   method = BCPlaid, name = "Plaid",
                                   transposed = TRUE)

# for questionnaire data - so haven't used BCQuest
bcspectral_results <- biclust_results(tc_data, docs_labs,
                                      method = BCSpectral, name = "spectral",
                                      transposed = TRUE)bcplaid_results2 <- biclust_results(lapply(tc_data, t), docs_labs,
                                    method = BCPlaid, name = "Plaid_t")
bcspectral_results2 <- biclust_results(lapply(tc_data, t), docs_labs,
                                       method = BCSpectral, name = "spectral_t")

# biclust(three_cancers[[2]], method = BCPlaid)

full_res <- rbind(bcplaid_results,
                  bcspectral_results)
# write.csv(full_res, "RealData/3cancers/biclust_results_term_doc.csv",
#           row.names = FALSE)
best_res <- matrix(0, nrow = 2, ncol = ncol(full_res))
i <- 1
for (method in c("Plaid", "spectral")) {
  res <- full_res[full_res[, 1] == method, ]
  best_res[i, ] <- res[which.max(res[, 5]), ]
  i <- i + 1
}
# only keep non transposed results
write.csv(best_res[c(1, 3), ], "RealData/3cancers/biclust_results.csv",
          row.names = FALSE)

cols <- apply(mvrnorm(n = 20267, mu = rep(0, 3), Sigma = diag(3)) > 0, 2, as.numeric)
bisilhouette(t(tc_data[[1]]),
        labels,
        cols,
        method = "cosine",
      )
```

Now look at the GFA results
```{r}
gfa_res <- gfa_apply(lapply(tc_data, t), dim(labels)[2])
colSums(gfa_res$col_clusters[[2]])
jacc_res <- all_jaccs(labels, gfa_res$row_clusters)
bisils <- calc_euc_sils(lapply(tc_data, t), gfa_res)
bisils <- calc_all_sils(lapply(tc_data, t), gfa_res)
gfa_full_res <- c("GFA", jacc_res, bisils)

write.csv(rbind(gfa_full_res,
                best_res),
          "RealData/3cancers/biclust_results.csv",
          row.names = FALSE)
```

## 
## 3cancers results
```{r}
source("SimStudy/Functions/extra_funcs.r")
source("SimStudy/Functions/evaluation_funcs.r")


three_cancers <- import_matrix("RealData/3cancers/tgca_3cancers.xlsx")
labels <- read.csv("RealData/3cancers/tgca_3cancers_labels.csv")
tc_data <- lapply(three_cancers, function(x) apply(x, 2, as.numeric))
tc_data <- lapply(tc_data, function(x) x[, 2:1242])
```

```{r}
library(biclust)
# no params needed
bcplaid_results <- biclust_results(tc_data, labels,
                                   method = BCPlaid, name = "Plaid",
                                   transposed = TRUE)

# for questionnaire data - so haven't used BCQuest
bcspectral_results <- biclust_results(tc_data, docs_labs,
                                      method = BCSpectral, name = "spectral",
                                      transposed = TRUE)bcplaid_results2 <- biclust_results(lapply(tc_data, t), docs_labs,
                                    method = BCPlaid, name = "Plaid_t")
bcspectral_results2 <- biclust_results(lapply(tc_data, t), docs_labs,
                                       method = BCSpectral, name = "spectral_t")

# biclust(three_cancers[[2]], method = BCPlaid)

full_res <- rbind(bcplaid_results,
                  bcspectral_results)
# write.csv(full_res, "RealData/3cancers/biclust_results_term_doc.csv",
#           row.names = FALSE)
best_res <- matrix(0, nrow = 2, ncol = ncol(full_res))
i <- 1
for (method in c("Plaid", "spectral")) {
  res <- full_res[full_res[, 1] == method, ]
  best_res[i, ] <- res[which.max(res[, 5]), ]
  i <- i + 1
}
# only keep non transposed results
write.csv(best_res[c(1, 3), ], "RealData/3cancers/biclust_results.csv",
          row.names = FALSE)

cols <- apply(mvrnorm(n = 20267, mu = rep(0, 3), Sigma = diag(3)) > 0, 2, as.numeric)
bisilhouette(t(tc_data[[1]]),
        labels,
        cols,
        method = "cosine",
      )
```

Now look at the GFA results
```{r}
gfa_res <- gfa_apply(lapply(tc_data, t), dim(labels)[2])
colSums(gfa_res$col_clusters[[2]])
jacc_res <- all_jaccs(labels, gfa_res$row_clusters)
bisils <- calc_euc_sils(lapply(tc_data, t), gfa_res)
bisils <- calc_all_sils(lapply(tc_data, t), gfa_res)
gfa_full_res <- c("GFA", jacc_res, bisils)

write.csv(rbind(gfa_full_res,
                best_res),
          "RealData/3cancers/biclust_results.csv",
          row.names = FALSE)
```

```{r}




bccc_res <- apply_biclust_method(lapply(three_data, t), BCCC, transposed=TRUE)
res_bccc <- biclust(t(three_data[[1]]), method = BCCC)
```


```{r}

results <- apply_biclust_method(three_data, method = BCQuest, transposed = TRUE)
dim(results$col_clusters[[1]])



dim(results$col_clusters[[3]])

bisils <- calc_all_sils(lapply(three_data, t), list(labels, labels, labels))
bisils_og <- calc_euc_sils(lapply(three_data, t), results)
bisils <- calc_euc_sils(three_data, list(row_clusters = results$col_clusters,
                                         col_clusters = results$row_clusters))

i <- 1
bisilhouette(t(three_data[[i]]),
        docs_labs,
        results$row_clusters[[i]],
        method = "euclidean",
      )$bisil

gfa_res <- gfa_apply(lapply(three_data, t), dim(docs_labs)[2])
gfa_res_t <- list(row_clusters = gfa_res$col_clusters,
                     col_clusters = gfa_res$row_clusters)

bisils_gfa <- calc_all_sils(lapply(three_data, t), gfa_res)
bisils_gfa_t <- calc_all_sils(three_data, gfa_res)

all_jaccs(docs_labs, gfa_res$row_clusters)

# assess performance
bisils <- calc_all_sils(lapply(three_data, t), bbc_res2[[k]])
jaccs <- all_jaccs(docs_labs, bbc_res2[[k]]$row_clusters)


all_jaccs(docs_labs,results$col_clusters)


jaccard_row(results$col_clusters[[3]], docs_labs)


res <- biclust(three_data[[2]], method = BCPlaid)
cols <- apply(res@NumberxCol, 2, as.numeric)
```

```{r}

all_res_3sources <- read.csv("RealData/3sources/all_results.csv")
issvd_gfa <- all_res_3sources[c(5, 6), ]
```